Target Speaker Extraction using Discrete Representations from Self-Supervised Models and Language Models \thanks{Identify applicable funding agency here. If none, delete this.}

### 2024-09-03 src/paper.tex
update

Discretization of audio has been studied due to the advancement of language models (LM). 
This approach translates the audio into discrete tokens simulating text vocabularies and uses LMs to 
model them. This approach simplifies audio generation tasks by transforming the complex 
regression problems into classification problems \cite{dasb}. There are currently two approaches for
audio discretization, the first one uses neural audio codecs \cite{dac}. This approach 
typically captures the acoustic features of the audio \cite{speech_tokenizer}. The second approach 
utilizes  self-supervised Learning (SSL) models like HuBERT \cite{hubert} and WavLM \cite{wavlm}.
SSL models have demonstrated 
excellent performances on many downstream tasks \cite{superb}. These SSL models extract  
continuous representations containing rich semantic and timbre information of a given speech. 
In this paper, we mainly explore how far we can go

>>>>

Discretization of audio has been studied due to the advancement of language models (LM). 
This approach translates the audio into discrete tokens simulating text vocabularies and uses LMs to 
model them. This approach simplifies audio generation tasks by transforming the complex 
regression problems into classification problems \cite{dasb}. There are currently two approaches for
audio discretization, the first one uses neural audio codecs \cite{dac}. This approach 
typically captures the acoustic features of the audio \cite{speech_tokenizer}. The second approach 
utilizes  self-supervised Learning (SSL) models like HuBERT \cite{hubert} and WavLM \cite{wavlm}.
SSL models have demonstrated 
excellent performances on many downstream tasks \cite{superb}. These SSL models extract  
continuous representations containing rich semantic and timbre information from a given speech. 
As demonstrated in \cite{dasb}, SSL models perform better than audio codec in tasks like speech 
enhancement and speech separation. Therefore, in this 
paper, we mainly explore the discretization of SSL models.

Discretization methods have been studied for speech enhancement \cite{selm} and blind speech separation \cite{dasb}, however, this approach has been rarely for target speaker extraction. TokenSplit \cite{tokensplit} needs the transcript of the speech to achieve satisfying results. In 
real life, however, speech is rarely provided with its transcript.
